A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2013-09-05 jacobian_calc_CMAME.tex
Adding figures
Cleaned up repo
Fix up first figure
Added results figures, finalized some removes

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

@@

\begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
matrix calculation methods including a newly developed Complex Taylor Series Expansion
(CTSE)-based ``complex-step'' method alongside established methods: forward-difference,
central-difference and automatic differentiation.  Tangent-stiffness matrix is a term from
computational mechanics which here refers to the Jacobian matrix as used in first-order
Newton-Raphson methods for the solution of non-linear algebraic systems of equations developed
to describe physical or theoretical systems studied in math, science, engineering and business.
To perform the comparative study, the above tangent-stiffness calculation methods were applied
in a massively parallel computational peridynamics code developed at Sandia National Labs,
called \emph{Peridigm}. In the comparative study, for each datum for each run, complex-step was
multiple orders of magnitude more accurate the finite-difference methods, according to a common
comparison made to the automatic-differentiation method datum. However, for the implementation
contemporary to the study, complex-step was also the slowest method among the four for each
datum for each run. A mathematical definition of \change[Michael]{the methods}{complex-step and
and automatic differentiation}, description of the computer implementations and justification
of the comparative study precede results for clarity. The intended audience of this paper
includes researchers, professionals and students, \change[Michael]{therefore conceptually explicit language
appealing to multiple levels of understanding is used}{therefor readers are directed to
selected background sources in the text for well-known concepts while less familiar concepts
are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
or who are new to their studies.}  \end{abstract}

@@

This subsection contains background information on the differentiation techniques underlying the
tangent-stiffness matrix calculation methods compared in the study. Since first-order finite
difference techniques are considered well known, they are not described where. Instead the
reader is referred to \cite[Chap. 4.1.3]{chapra2010}.

\subsubsection{Basis of New Method, 'Complex Step', in CTSE}  
\note[Michael]{TODO: cite lyness and moler, complete section}

@@

\remove[Michael]{
\subsubsection{First-order Finite-difference}

The types of first-order finite-difference that will be discussed are forward-difference and
central-difference. CTSE and these finite-difference techniques are similar in that they have been
developed by algebraically solving approximated Taylor series expansions for first derivative terms.
}

@@

\subsubsection{Solve a Nonlinear System with a Tangent-stiffness Matrix} \label{solve_a_system}
\remove[Michael]{ 
An accurate tangent-stiffness matrix is a main part of the Newton-Raphson method for iteratively
solving non-linear systems for equilibrium solutions. The method is derived by a linearisation using
Taylor expansion of the vector valued function modelling the system in a way reminiscent of how
forward-difference is derived from the Taylor expansion of a scalar valued function. Here the two
derivations are shown side by side: The difference between the two is that in the forward difference
derivation, it is assumed that the derivative term is unknown and can be determined from knowing
change in location, while in the Newton-Raphson derivation, it is assumed that the change in
location is unknown and can be determined from knowing the value of the derivative term.
Additionally we could derive a version of Newton Raphson for CTSE directly: The Newton Raphson
method is applied in the following manner: It is clear to see how if the tangent-stiffness matrix
used in the Newton Raphson method is inaccurate, predicted changes in location will be inaccurate.
However, non-linear functions don't have constant slopes by definition, meaning that a perfectly
accurate tangent-stiffness matrix at the current guess location is not a guarantee of a one step
solution path to the solution location since slope changes over the interval guess to updated guess.
}
The reader is referred to a succinct explanation of the Newton Raphson method, which shows the
role of a tangent-stiffness matrix or Jacobian in that solution method \cite[chap.
13]{young2009}.

@@

\subsection{Speed Data}
\subsection{Accuracy Data}
\subsection{Efficiency Data}

>>>>

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%%\usepackage{graphics}
%% or use the graphicx package for more complicated commands
%%\usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}
%%\usepackage{color}
%%\usepackage{tkz-base}
\usepackage{pgfplots}

@@

\begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
matrix calculation methods including a newly developed Complex Taylor Series Expansion
(CTSE)-based ``complex-step'' method alongside established methods: forward-difference,
central-difference and automatic differentiation.  Tangent-stiffness matrix is a term from
computational mechanics which here refers to the Jacobian matrix as used in first-order
Newton-Raphson methods for the solution of non-linear algebraic systems of equations developed
to describe physical or theoretical systems studied in math, science, engineering and business.
To perform the comparative study, the above tangent-stiffness calculation methods were applied
in a massively parallel computational peridynamics code developed at Sandia National Labs,
called \emph{Peridigm}. In the comparative study, for each datum for each run, complex-step was
multiple orders of magnitude more accurate the finite-difference methods, according to a common
comparison made to the automatic-differentiation method datum. However, for the implementation
contemporary to the study, complex-step was also the slowest method among the four for each
datum for each run. A mathematical definition of complex-step and
and automatic differentiation, description of the computer implementations and justification
of the comparative study precede results for clarity. The intended audience of this paper
includes researchers, professionals and students, therefor readers are directed to
selected background sources in the text for well-known concepts while less familiar concepts
are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
or who are new to their studies.  \end{abstract}

@@

This subsection contains background information on the differentiation techniques underlying the
tangent-stiffness matrix calculation methods compared in the study. Since first-order finite
difference techniques are considered well known, they are not described here. Instead the
reader is referred to \cite[Chap. 4.1.3]{chapra2010}.

\subsubsection{Basis of New Method, 'Complex Step', in CTSE}  
\note[Michael]{Will be replaced by Dr. Millwater's background section}

@@

@@

\subsubsection{Solve a Nonlinear System with a Tangent-stiffness Matrix} \label{solve_a_system}
The reader is referred to a succinct explanation of the Newton Raphson method, which shows the
role of a tangent-stiffness matrix or Jacobian in that solution method \cite[chap.
13]{young2009}.

@@

\subsection{Speed Data}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_speed.pgf}}
  \caption{Serial test series speed measurements.}
  \label{fig:serial_speed}
\end{figure}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_speed.pgf}}
  \caption{Multicore test series speed measurements.}
  \label{fig:multi_speed}
\end{figure}

\subsection{Accuracy Data}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_accuracy.pgf}}
  \caption{Serial test series speed measurements.}
  \label{fig:serial_accuracy}
\end{figure}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_accuracy.pgf}}
  \caption{Multicore test series accuracy measurements.}
  \label{fig:multi_accuracy}
\end{figure}

\subsection{Efficiency Data}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_speed_rel.pgf}}
  \caption{Serial test series efficiency measurements.}
  \label{fig:serial_efficiency}
\end{figure}

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_speed_rel.pgf}}
  \caption{Multicore test series efficiency measurements.}
  \label{fig:multi_efficiency}
\end{figure}

%%\includegraphics[width=\textwidth]{serial_speed_rel.png}
%%\includegraphics[width=\textwidth]{multi_speed_rel.png}
